---
title: "Interactive charts using Shiny"
author: "Sneh Bindesh Chitalia"
date: "20/06/2020"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Detail about the charts

mtcars dataset has been used for this shiny web application. The outcome is fixed which is Miles Per Gallon (mpg), whereas the x-axis variable can be changed, furthermore, the variables controlling color and size of each data point can also be changed. The web application is hosted on https://csneh4.shinyapps.io/Interactive_Prediction/ .  
The code for this is available on https://github.com/sneh4306/Shiny-Web-Application

## About the Data

```{r echo=TRUE}
head(mtcars)
```

## UI code

```{r, echo = TRUE, eval=FALSE}
library(shiny)

# Define UI for application that draws a histogram
shinyUI(fluidPage(

    # Application title
    titlePanel("Cars Dataset"),

    # Sidebar with a slider input for number of bins
    sidebarLayout(
        sidebarPanel(
        selectInput("xvar","Variable for X-axis",
                    choices=list("Cylinder"="cyl","Transmission"="am","Gears"="gear"),selected = "cyl"),
        selectInput("color","Variable for color",
                    choices=list("Carburetors"="carb","Engine Type"="vs"),selected="carb"),
        selectInput("size","Variable for size",
                    choices=list("Horsepower"="hp","Weight"="wt"),selected = "hp"),
        submitButton("Submit")
        ),
    
    
    

        # Show a plot of the generated distribution
        mainPanel(
            h3(textOutput("Title")),
            plotOutput("plot")
        )
    )
    ))


```

## Server Code

```{r,echo=TRUE,eval=FALSE}
library(shiny)
library(ggplot2)


# Define server logic required to draw a histogram
shinyServer(function(input, output) {


    mtcars$vs<-factor(mtcars$vs,labels = c("V-shaped","Straight"))
    mtcars$carb<-as.factor(mtcars$carb)
    text<-reactive({
        paste("mpg vs",input$xvar)
    })
    output$Title<-renderText({
        text()
        })
    
    p<-reactive({
        ax<-list(
            title=input$xvar
        )
        al<-list(
            title=input$color
        )
        
        g<-ggplot(mtcars,aes_string(y="mpg",x=input$xvar,col=input$color))+geom_point(aes_string(size=input$size))
        g
        })
   
    output$plot<-renderPlot({
        p()
         })
})

```

