---
title: "Basic Differential Data Analysis"
author: "Sneh Bindesh Chitalia"
date: "11/06/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the dataset
```{r}
library(datasets)
data("ToothGrowth")
head(ToothGrowth)
```

## Basic Exploratory Data Analysis
```{r}
library(ggplot2)
#Plot 1 (dose vs len)
ToothGrowth$dose<-as.factor(ToothGrowth$dose)
g<-ggplot(ToothGrowth,aes(x=dose,y=len,fill=dose))+geom_boxplot()+ggtitle("Comparison between dosage and tooth length")+xlab("dose in mg")+ylab("tooth length")
print(g)
#Plot 2 (supp vs len)
k<-ggplot(ToothGrowth,aes(x=supp,y=len,fill=supp))+geom_boxplot()+ggtitle("Comparison of delivery method with tooth length")+xlab("delivery method")+ylab("tooth length")
print(k)
```

## Summary  
Check for any NA or Null or any type of missing values
```{r}
sum(!complete.cases(ToothGrowth))
```

Thus, there are no missing values in the data.  
Further exploring the data
```{r}
summary(ToothGrowth)
```

This gives us an overview of the data
```{r}
str(ToothGrowth)
```

This gives us information about data type of each variable.  
  
## Hypothesis Testing  
We carry out hypothesis testing on whether delivery mode has any impact on teeth length.  
The null hypothesis is:  
# H0: The delivery mode for Vitamin C does not have any impact on tooth length
```{r}
dosage<-ToothGrowth$dose
supp<-ToothGrowth$supp
len<-ToothGrowth$len
t.test(len[supp=="OJ"],len[supp=="VC"],paired=FALSE)
```

Since the p-value > 0.05, we fail to reject the null hypothesis.  
However, the confidence interval contains 0 which indicates that the test is not significant.  
  
# H0: The dosage in mg does not have any impact on tooth length
```{r}
# compare between dosage of 0.5 mg and 1 mg
t.test(len[dosage==0.5],len[dosage==1],paired = FALSE)
```

Since the p-value < 0.05, we can reject the null hypothesis.  
Also, the confidence interval does not contain zero which indicates that the test is significant.
```{r}
# compare between dosage of 1 mg and 2 mg
t.test(len[dosage==1],len[dosage==2],paired=FALSE)
```

Since the p-value < 0.05, we can reject the null hypothesis.  
Also, the confidence interval does not contain zero which indicates that the test is significant.  
Thus we can succesfully rule out null hypothesis.  
  
## Conclusions   
1. The delivery method of Vitamin C does not impact the teeth growth.  
2. The amount of dosage given to the guinea pigs impact the teeth growth.    
  
## Assumptions
1. t-tests are considered.    
2. These t-tests are considered to be unpaired.