---
title: "Activity Monitoring"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```
## Loading and preprocessing the data
Read Data
```{r}
data<-read.csv("activity.csv")
head(data)
```

Histogram of the total number of steps taken each day
```{r}
library(dplyr)
library(ggplot2)
k<-data %>% select(date,steps) %>% group_by(date) %>% summarise(Total_steps=sum(steps))%>% na.omit()
hist(k$Total_steps,breaks=25,xlab="Total Steps",main="Total Steps by day",col="green")
```

## What is mean total number of steps taken per day?
Mean number of steps taken each day
```{r}
mean(k$Total_steps)
```
Median number of steps taken each day
```{r}
median(k$Total_steps)
```
## What is the average daily activity pattern?
Time series plot of the average number of steps taken
```{r}
z<-data %>% select(interval,steps) %>% na.omit() %>% group_by(interval) %>% summarise(Total_steps=mean(steps))
g<-ggplot(z,aes(interval,Total_steps))+geom_line()+ggtitle("Time series plot of average number of steps taken")
print(g)
```
The 5-minute interval that, on average, contains the maximum number of steps
```{r}
z[which(z$Total_steps==max(z$Total_steps)),]
```
## Imputing missing values
Code to describe and show a strategy for imputing missing data
Number of missing values in the data(according to date)
```{r}
sum(is.na(data$date))
```
Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
```{r}
replace_mean<-function(x){replace(x,is.na(x),mean(x,na.rm=TRUE))}
data_mean<-data %>% group_by(interval) %>% mutate(steps=replace_mean(steps))
head(data_mean)
```
Histogram of the total number of steps taken each day after missing values are imputed
```{r}
dat<-data_mean %>% group_by(date) %>% summarise(Total_steps=sum(steps))
hist(dat$Total_steps,breaks=25,xlab="Total Steps",main="Total Steps by day",col="green")
```
Comparison between old and new mean and median
```{r}
oldmean <- mean(k$Total_steps, na.rm = TRUE)
newmean <- mean(dat$Total_steps)
# Old mean and New mean
oldmean
newmean
oldmedian <- median(k$Total_steps, na.rm = TRUE)
newmedian <- median(dat$Total_steps)
# Old median and New median
oldmedian
newmedian
```
## Are there differences in activity patterns between weekdays and weekends?
Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
1. Convert date into days and label them as weekdays and weekends
```{r}
data_mean$date<-as.Date(as.character(data_mean$date))
data_mean$type<-ifelse(weekdays(data_mean$date)=="Saturday" | weekdays(data_mean$date)=="Sunday", "Weekend", "Weekday" )
```
2. Plot the graph
```{r}
data_type<-aggregate(data_mean$steps,by=list(data_mean$type,data_mean$interval),na.omit(mean))
names(data_type)=c("type","interval","steps")
ggplot(data_type, aes(x=interval, y=steps, color=type)) + geom_line()+
     facet_grid(type ~.) + xlab("Interval") + ylab("Mean of Steps") +
     ggtitle("Comparison of Average Number of Steps in Each Interval")
```